
---
title: "Group Project Notebook"
output: html_notebook
---

1. Libraries & Data Access

```{r   }
library(data.table)   
library(ggplot2)
library(igraph)

dt.books <- fread("dt.books.clean.csv") 
save(dt.books, file="books.RData") 
```

2. Loading the dataset

```{r   }
load("books.RData") 

# Creating the column avg_rating_class that tells for each BOOK the rating that they are in
dt.books$avg_rating_class <- with(dt.books, ifelse(average_rating >= 0 & average_rating < 0.5, "[0-0.5[", ifelse(average_rating >= 0.5 & average_rating < 1, "[0.5 -1[", ifelse(average_rating >= 1 & average_rating < 1.5, "[1-1.5[", ifelse(average_rating >= 1.5 & average_rating < 2, "[1.5-2[", ifelse(average_rating >= 2 & average_rating <= 2.5, "[2-2.5]", ifelse(average_rating >= 2.5 & average_rating <= 3, "[2.5-3[", ifelse(average_rating >= 3 & average_rating <= 3.5, "[3-3.5[", ifelse(average_rating >= 3.5 & average_rating <= 4,  "[3.5-4[", ifelse(average_rating >= 4 & average_rating <= 4.5,  "[4-4.5[", ifelse(average_rating >= 4.5 & average_rating <= 5,  "[4.5-5]", "NA")))))))))))

# Creating the column avg_rating_individual that tells for each AUTHOR the rating that they are in (by doing the average of all their books)
average.ratings <- vector()

for (author in dt.books$authors) {
  the_author <- dt.books[authors %in% author]
  average_rating_author <- mean(the_author$average_rating)
  average.ratings <- c(average.ratings, average_rating_author)
}

dt.books$avg_rating_individual <- average.ratings

# Creating the column avg_rating_individual_class that tells for each AUTHOR the rating that they are in (same as above but with new column)
dt.books$avg_rating_individual_class <- with(dt.books, ifelse(avg_rating_individual >= 0 & avg_rating_individual < 0.5, "[0-0.5[", ifelse(avg_rating_individual >= 0.5 & avg_rating_individual < 1, "[0.5 -1[", ifelse(avg_rating_individual >= 1 & avg_rating_individual < 1.5, "[1-1.5[", ifelse(avg_rating_individual >= 1.5 & avg_rating_individual < 2, "[1.5-2[", ifelse(avg_rating_individual >= 2 & avg_rating_individual <= 2.5, "[2-2.5]", ifelse(avg_rating_individual >= 2.5 & avg_rating_individual <= 3, "[2.5-3[", ifelse(avg_rating_individual >= 3 & avg_rating_individual <= 3.5, "[3-3.5[", ifelse(avg_rating_individual >= 3.5 & avg_rating_individual <= 4,  "[3.5-4[", ifelse(avg_rating_individual >= 4 & avg_rating_individual <= 4.5,  "[4-4.5[", ifelse(avg_rating_individual >= 4.5 & avg_rating_individual <= 5,  "[4.5-5]", "NA")))))))))))
```

3. Choosing the author

```{r   }
author = "Alan Dean Foster"
```

4. Average ranking of the author

```{r   }
dt.author <- dt.books[authors == author]
dt.author$avg_rating_individual[1]
```


5. Number of books published

```{r   }
nrow(dt.author)
```

6. Most common category of the author

```{r   }
categories.count <- as.data.frame(table(dt.author$categories)) 
categories.count <- categories.count[order(-categories.count$Freq),]
categories.count$Var1[[1]]
```

7. With how many co-authors has the author colaborated with?

```{r   }

```

8. Graph that connects author with the book

```{r   }
dt.author.unique <- dt.author[!duplicated(dt.author[,title])]
dt.author.unique <- dt.author.unique[,c("title", "authors")]
g.author.books <- graph.data.frame(dt.author.unique, directed=TRUE)
plot(g.author.books)
```

9. Graph that connects books to the category/categories

```{r   }
dt.author.category <- dt.author.unique[,c("title", "categories")]
g.books.category <- graph.data.frame(dt.author.category, directed=TRUE, vertices = NULL)
plot(g.books.category)
```

10. Graph that links to other authors that have a similar ranking

```{r   }

#Defined by user
N = 30

author_class <- dt.author$avg_rating_individual_class[1]

dt.books.unique <- dt.books[!duplicated(dt.books[,authors])]
dt.books.unique.similar <- dt.books.unique[avg_rating_individual_class == author_class]
dt.books.unique.similar <- dt.books.unique.similar[,c("authors", "avg_rating_individual_class")]
g.books.ranking <- graph.data.frame(dt.books.unique.similar[1:N], directed=TRUE, vertices = NULL)
plot(g.books.ranking)
```

11. Graph that links authors with the same category

```{r   }
```

12. Graph that links with the same author and category

```{r   }
```

13. Graph that links with authors that published on the same year

```{r   }
```