
---
title: "Group Project Notebook"
output: html_notebook
---

1. Libraries & Data Access

```{r   }
library(data.table)   
library(ggplot2)
library(igraph)
library(openxlsx)

dt.books <- read.xlsx("dt.books.clean.xlsx")
save(dt.books, file="books.RData") 
```

2. Loading the dataset

```{r   }
load("books.RData") 
```

2.1 Creating columns: avg_rating_class, avg_rating_individual & avg_rating_individual_cass

```{r   }
dt.books <- na.omit(dt.books)

# Creating the column avg_rating_class that tells for each BOOK the rating that they are in
dt.books$avg_rating_class <- with(dt.books, ifelse(average_rating >= 0 & average_rating < 0.5, "[0-0.5[", ifelse(average_rating >= 0.5 & average_rating < 1, "[0.5 -1[", ifelse(average_rating >= 1 & average_rating < 1.5, "[1-1.5[", ifelse(average_rating >= 1.5 & average_rating < 2, "[1.5-2[", ifelse(average_rating >= 2 & average_rating <= 2.5, "[2-2.5]", ifelse(average_rating >= 2.5 & average_rating <= 3, "[2.5-3[", ifelse(average_rating >= 3 & average_rating <= 3.5, "[3-3.5[", ifelse(average_rating >= 3.5 & average_rating <= 4,  "[3.5-4[", ifelse(average_rating >= 4 & average_rating <= 4.5,  "[4-4.5[", ifelse(average_rating >= 4.5 & average_rating <= 5,  "[4.5-5]", "NA")))))))))))

# Creating the column avg_rating_individual that tells for each AUTHOR the rating that they are in (by doing the average of all their books)
average.ratings <- vector()

for (author in dt.books$authors) {
  the_author <- dt.books[dt.books$authors %in% author, ]
  average_rating_author <- mean(the_author$average_rating)
  average.ratings <- c(average.ratings, average_rating_author)
}

dt.books$avg_rating_individual <- average.ratings

# Creating the column avg_rating_individual_class that tells for each AUTHOR the rating that they are in (same as above but with new column)
dt.books$avg_rating_individual_class <- with(dt.books, ifelse(avg_rating_individual >= 0 & avg_rating_individual < 0.5, "[0-0.5[", ifelse(avg_rating_individual >= 0.5 & avg_rating_individual < 1, "[0.5 -1[", ifelse(avg_rating_individual >= 1 & avg_rating_individual < 1.5, "[1-1.5[", ifelse(avg_rating_individual >= 1.5 & avg_rating_individual < 2, "[1.5-2[", ifelse(avg_rating_individual >= 2 & avg_rating_individual <= 2.5, "[2-2.5[", ifelse(avg_rating_individual >= 2.5 & avg_rating_individual <= 3, "[2.5-3[", ifelse(avg_rating_individual >= 3 & avg_rating_individual <= 3.5, "[3-3.5[", ifelse(avg_rating_individual >= 3.5 & avg_rating_individual <= 4,  "[3.5-4[", ifelse(avg_rating_individual >= 4 & avg_rating_individual <= 4.5,  "[4-4.5[", ifelse(avg_rating_individual >= 4.5 & avg_rating_individual <= 5,  "[4.5-5]", "NA")))))))))))
```

2.2 Creating column main_category: 
```{r   }
main.categories <- vector()

for (author in dt.books$authors) {
  the_author <- dt.books[dt.books$authors %in% author, ]
  main.category <- names(sort(table(the_author$categories), decreasing=TRUE)[1])
  main.categories <- c(main.categories, main.category)
}
dt.books$main_category <- main.categories
```

2.3 Creating columns n_books, n_books_class
```{r   }
n.books <- vector()

for (author in dt.books$authors) {
  the_author <- dt.books[dt.books$authors %in% author, ]
  #the.author.unique <- the_author[!duplicated(the_author$title), ]
  n.books.author <- nrow(the_author)
  n.books <- c(n.books, n.books.author)
}
dt.books$n_books <- n.books

# Creating the column n_books_class that tells for each AUTHOR the class of number of books they are in
dt.books$n_books_class <- with(dt.books, ifelse(n.books >= 0 & n.books < 2, "[0-2[", ifelse(n.books >= 2 & n.books < 5, "[2-5[", ifelse(n.books >= 5 & n.books < 10, "[5 -10[", ifelse(n.books >= 10 & n.books < 15, "[10-15[", ifelse(n.books >= 15 & n.books < 20, "[15-20[", ifelse(n.books >= 20 & n.books <= 25, "[20-25[", ifelse(n.books >= 25 & n.books <= 30, "[25-30[", ifelse(n.books >= 30 & n.books <= 35, "[30-35[", ifelse(n.books >= 35 & n.books <= 40, "[35-40[", ifelse(n.books >= 40 & n.books <= 45, "[40-45]", "NA")))))))))))
```


3. Choosing the author

```{r   }
author = "Margaret Weis"
```

4. Average ranking of the author

```{r   }
dt.author <- dt.books[dt.books$authors %in% author, ]
dt.author$avg_rating_individual[1]
```



5. Number of books published

```{r   }
nrow(dt.author)
```


6. Years where the actor published books
```{r   }
dt.unique.years <- unique(dt.author$published_year)
sort(dt.unique.years, decreasing = FALSE)
```



7. Most common category of the author

```{r   }
dt.author$main_category[1]
```


8. With how many co-authors has the author colaborated with?

```{r   }
author.books <- unique(dt.author$title)
books.in.dt <- dt.books[dt.books$title %in% author.books, ]
books.in.dt.other <- books.in.dt[!books.in.dt$authors %in% author, ]
books.in.dt.other[,c("title", "authors")]
```

9. Graph that connects author with the book

```{r   }
dt.author.unique <- dt.author[!duplicated(dt.author$title), ]
dt.author.books <- dt.author.unique[,c("title", "authors")]
g.author.books <- graph.data.frame(dt.author.books, directed=FALSE)
plot(g.author.books)
```

10. Graph that connects books to the category/categories

```{r   }
dt.author.category <- dt.author.unique[,c("title", "categories")]
g.books.category <- graph.data.frame(dt.author.category, directed=FALSE, vertices = NULL)
plot(g.books.category)
```

11. Graph that links to other authors that have a similar ranking and a similar nº of books published

```{r   }
author_class <- dt.author$avg_rating_individual_class[1]
author_n_books_class <- dt.author$n_books_class[1]

dt.books.unique <- dt.books[!duplicated(dt.author$authors), ]
dt.books.unique.similar.rank <- dt.books.unique[dt.books.unique$avg_rating_individual_class == author_class & dt.books.unique$n_books_class == author_n_books_class, ]

dt.books.unique.similar.rank$main.author <- author
dt.books.unique.similar.rank <- dt.books.unique.similar.rank[,c("main.author", "authors")]
dt.books.unique.similar.rank <- dt.books.unique.similar.rank <- dt.books.unique.similar.rank[!duplicated(dt.books.unique.similar.rank$authors), ]

g.books.ranking <- graph.data.frame(dt.books.unique.similar.rank, directed=FALSE, vertices = NULL)
plot(g.books.ranking)
```

12. Graph that links authors with the same main category

```{r   }
author_main_category <- dt.author$main_category[1]

dt.books.unique.category <- dt.books.unique[dt.books.unique$main_category == author_main_category & dt.books.unique$n_books_class == author_n_books_class,  ]

dt.books.unique.category$main.author <- author
dt.books.unique.category <- dt.books.unique.category[,c("main.author", "authors")]
g.books.same.category <- graph.data.frame(dt.books.unique.category, directed=FALSE, vertices = NULL)
plot(g.books.same.category)
```


13. Graph that links with similar ranking AND category (10 and 11 combined)

```{r   }
dt.merged <- merge(dt.books.unique.similar.rank, dt.books.unique.category ,by="authors", all = FALSE)

dt.merged$main.author <- author
dt.merged <- dt.merged[,c("main.author", "authors")]
  
g.merged <- graph.data.frame(dt.merged, directed=FALSE, vertices = NULL)
plot(g.merged)
```